vary_distribution

### 2020-12-18 vary_distribution.Rmd
all raw results after transformation added

---
title: Selecting and ranking interesting pairs of cyclic temporal granularities
authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  email: Sayani.Gupta@monash.edu

bibliography: bibliography.bib
preamble: >
  \usepackage{mathtools,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \usepackage{algorithm}
  \usepackage{algorithmicx}
  \usepackage{algpseudocode}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE,
        tinytex.verbose = FALSE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(viridis)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)

opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
#                                                                read_chunk('main.R')
here::here()
read_chunk(here::here("paper/R/", "null_distribution.R"))
read_chunk(here::here("paper/R/", "mean_null.R"))
source(here::here("paper/R/sim_panel.R"))
# Set up plan
#source("_drake.R")
# Run all code required
#drake::r_make()
# load each object as required
# loadd()
```

```{r}
set.seed(321)
```

# Introduction

<!-- <introducing the problem> -->
<!-- background of the problem -->

Exploratory data analysis, as coined by John W. Tukey (Tukey 1965) involves many iterations of finding structures and patterns that allows the data to be informative. With temporal data available at finer scales, exploring periodicity and their relationships can become overwhelming with so many possible cyclic temporal granularities [@Gupta2020-vo] to explore.

```{r calendar-elec, fig.height = 10.5, fig.cap="Calendar display."}
elec <- read_rds(here("paper/data/elec.rds")) %>% 
  filter(date >= ymd("20180101"), date < ymd("20180701"))
rdbl <- c("Weekday" = "#d7191c", "Weekend" = "#2c7bb6")

elec <- elec %>% 
  mutate(
    wday = wday(date, label = TRUE, week_start = 1),
    weekday = if_else(wday %in% c("Sat", "Sun"), "Weekend", "Weekday")
  )
p_cal_elec <- elec %>% 
  filter(id %in% c(2, 4)) %>% 
  frame_calendar(x = time, y = kwh, date = date, nrow = 1) %>% 
    ggplot(aes(x = .time, y = .kwh, group = date)) +
    geom_line(aes(colour = as.factor(id)), size = 0.5) +
    scale_colour_brewer(name = "", palette = "Dark2", direction = 1) +
    facet_grid(id ~ ., labeller = label_both) +
    theme(legend.position = "bottom")
prettify(p_cal_elec, size = 2.5, label.padding = unit(0.1, "lines"))
```

\noindent Take the example of the calendar display of electricity smart meter data (\ref{fig:calendar-elec}) used in @wang2020calendar for four households in Melbourne, Australia. The authors show how hour-of-the-day interact with weekday and weekends and then move on to use calendar display to show daily schedules. The calendar display has several components in it, which helps us look at energy consumption across hour-of-the-day, day-of-the-week, week-of-the-month, and month-of-the-year at once. Some interaction of these cyclic granularities could also be interpreted from this display. This is a great start to have an overview of the energy consumption. However, if one wants to understand the periodicities in energy behavior and how the periodicities interact in greater details, it is not easy to comprehend the interactions of some periodicities' from this display, due to the combination of linear and cyclic representation of time. For example, this display might not be the best to understand how hour-of-the-day varies and month-of-year varies across week-of-the-month. Further, it is not clear what all interactions of cyclic granularities should be read from this display as there could be many combinations that one can look at. Moreover, calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions (@Gupta2020-vo) and could include other cyclic granularities like hour-of-week or day-of-fortnight, which could potentially become useful depending on the context.

```{r intro_all}

id2_tsibble <- elec %>% 
   filter(id == 2) %>% 
   as_tsibble(index = date_time)

id4_tsibble <- elec %>% 
   filter(id == 4) %>% 
   as_tsibble(index = date_time)

# hour-of-day and month-of-year (important pair) id2's behavior across different hours of the day very different across months, but for id4 behavior across different hours of the day is not likely a function of month.
p1 <- id2_tsibble %>%
   prob_plot("month_year",
             "hour_day",
             response = "kwh",
             plot_type = "quantile",
             symmetric = TRUE,
             quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9)) +
  ggtitle("") + theme(
        strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + 
  scale_colour_brewer(name = "", palette = "PiYG")

p2 <- id4_tsibble %>%
   prob_plot("month_year",
             "hour_day",
             response = "kwh",
             plot_type = "quantile",
             symmetric = TRUE,
             quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9)) + 
  ggtitle("") + theme(
        strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

# wknd_wday and week_month (important pair) id2's behavior across different hours of the day very different across months, but for id4 behavior across different hours of the day is not likely a function of month.

p3 <- id2_tsibble %>%
  create_gran("week_month") %>% 
  filter(week_month != 5) %>% 
   prob_plot("wknd_wday",
             "week_month",
             response = "kwh",
             plot_type = "quantile",
             symmetric = FALSE,
             quantile_prob = c(0.25,0.5,0.75)) +
   ggtitle("") +
#scale_x_discrete(breaks = seq(0, 23, 4))  + theme(
        theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

p4 <- id4_tsibble %>%
    create_gran("week_month") %>% 
  filter(week_month != 5) %>% 
   prob_plot("wknd_wday",
             "week_month",
             response = "kwh",
             plot_type = "quantile",
             symmetric = FALSE,
             quantile_prob = c(0.25,0.5,0.75)) +
  ggtitle("") +  #+ scale_x_discrete(breaks = seq(0, 23, 4)) + 
theme(
        strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

```

```{r id2, fig.cap = "something"}
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r id4, fig.cap = "something2"}
ggpubr::ggarrange(p3, p4, ncol = 2)
```

Moreover, there might be specific interactions that are interesting and others that are not and that too will vary with different households. For example, area distribution quantiles are plotted for household 2 and 4 in Figure \ref{fig:id2}a and b respectively. For the first household, the 75th and 90th percentile for Jan, Feb and July are very close, implying that energy usage for these months are generally on a much higher side due to the usage of air conditioners (in Jan and Feb) and heaters (in July). The energy consumption for household 2 is also higher relative to its own consumption for Jan, Feb and March but the 75th and 90th percentile are apart implying that contrary to the first household, the second household resorts to air conditioners and heaters much less regularly than the first one. Moreover, the 75th percentile distribution is not bimodal across hours of the day for the first household in those months, but the distribution looks similar for all months for the second household. Difference in the energy consumption seem to be varying both across month-of-year (facets) and hour-of-day (x-axis). And thus, both the cyclic granularities would deem important while studying the periodicities in the first household. However, it seems like energy consumption across hours of the day are not that different across different months for the second household. Differences seem to be more prominent across month-of-year (facets) than hour-of-day (x-axis). Again, look at \ref{fig:} c and d, where energy consumption for these two households are plotted against (weekend/weekday, week-of-month). Here, for both households, the pattern of energy consumption vary  across different weeks of the month irrespective of the fact it is a weekday or weekend. In that respect, the harmony pair (month-of-year, hour-of-day) seems to be more informative than (weekend/weekday, week-of-month) for the first household. It could be immensely useful to make the transition from all possible ways to only ways that could potentially be informative given a household.

<!-- Take an example of a data set which are observed at fine temporal scales, like that of NYC bike usage available at https://www.citibikenyc.com/system-data. We use the `nyc_bikes` data set from the R package `tsibbledata` which takes a sample of 10 bikes for the year 2018. The `start_time` and the `stop_time` are recorded to a fineness of seconds. We can look at pair of cyclic granularities (hour_day, wknd_wday) or (week_month, day_week) to see how these periodicities interact. But there could be other pairs that are important too. How to understand which pairs are sufficient to explore given the data set without losing much information about the data. -->

<!-- When we need to understand the interplay of different periodicities in a high frequency temporal datasets, we have many choices to consider. In [@wang2019tsibble] and [@wang2020calendar], periodicities are explored across hour of the day and day of the week or months. But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions. -->

<!-- what is the dimension of the problem -->
The paper @Gupta2020-vo describes how we can compute all possible combinations of cyclic time granularities. If we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed. Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that can originate from the underlying periodic and aperiodic linear granularities. The mapping of the graphical elements chosen in the paper implies that, for a numeric response variable, the graphics display distributions across combinations of cyclic granularities, one placed at x-axis and the other on the facet. That essentially implies there are $^{N_C}P_2$ possible pairwise plots exhaustively, where each plot would display a pair of cyclic granularities. This is large and overwhelming for human consumption.

<!-- Scagnostics literature -->
This is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to discern meaningful patterns in large collections of scatterplots. Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots. Therefore
for even small $v$, the number of scatterplots can be
large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data. @Dang2014-tw and @wilkinson2005graph provides potential solutions to this, where few characterizations help us to locate anomalies for defining several measures aimed to detect anomalies in density, shape, trend, and
other features in the 2D point scatters.

<!-- harmonies and why it is not enough -->
The paper (@Gupta2020-vo) narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash"). However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration. Hence, there is a need to reduce the search even further by including only those harmonies which are informative enough. Also, ranking the remaining harmony pairs based on how well they capture the variation in the measured variable could be potentially useful.

In this paper, we aim to build a new measure to follow through these two main objectives:

- To choose harmonies for which distributions of categories are significantly different 
- To rank the selected harmonies from highest to lowest variation in the distribution of their categories.

<!-- Talk about Scagnostics: Tukey -->

# Distance measure

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Median Maximum Pairwise Distances (MMPD) to evaluate structure in such a design.

## Idea {#sec:idea}

The principle employed for building a new metric is explained through a simple example explained in Figure \ref{fig:null4by2}. Each of these figures have the same panel design with 2 x-axis categories and 4 facet levels. Figure \ref{fig:null4by2}a has all x categories drawn from N(5, 10) distribution for each facet. It is not an interesting display particularly, as distributions do not vary across x-axis or facet categories. Figure \ref{fig:null4by2}b has x categories drawn from the same distribution within a facet and different for different facet categories. Figure \ref{fig:null4by2}b exhibits an exact opposite situation where distribution between the x-axis categories within each facet is different but they are same across facets. Figure \ref{fig:null4by2}d takes a step further by varying the distribution across both facet and x-axis categories. If we are asked to rank the displays in order of importance from minimum to maximum, we might order it as a, b, c and then d. It might be argued that it is not clear if b should precede or succeed c. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data. With this principle, display b could be considered less informative as compared to display c.

With reference to the graphical design in \ref{@Gupta2020-vo}, therefore the idea would be to rate a harmony pair higher if the variation between different levels of the x-axis variable is higher on an average across all levels of the facet variables. Thus the metric could be obtained by computing maximum pairwise distances between distributions of the continuous random variable across x-axis categories for all facets and then taking the median of those maximum pairwise distances across facets. This would help capture the average maximum difference in distribution of the measurement variable explained by the two cyclic granularities together. We call this metric MMPD which stands for Median Maximum Pairwise Distances. In the next section we shall see how we go about computing this measure.

<!-- To elaborate further, look at the examples in Figure \ref{}, where Figure \ref{}a represents the panel design with distribution of each x categories drawn from N(5, 10) distribution. It could be observed that the graph is not particularly interesting, as there is no significant change in distribution between x-axis levels or facets. Figure \ref{}b represents the same panel design with no difference in distribution of x-axis categories within a facet, but different distribution of x-axis categories for different facets. For example, if there are 4 facet levels and 2 x-axis levels, data is generated in the way as described in Table \ref{}. Figure \ref{}b exhibits an exact opposite situation where the x-axis within facets are different but not across facets. -->
<!-- Figure \ref{}d takes it further by varying the distribution across both facet and x-axis categories. -->

```{r null4by2,fig.cap=" A graphical display with two categories mapped to x-axis and 4 categories mapped to facets with the distribution of a continuous random variable plotted on the y-axis. Display a is not interesting as the distribution of the continuous rv does not depend across x-axis or facet categories. Display b and c are more interesting than a since there is a change in distribution either across facets(b) or x-axis(a). Display d is most interesting as distribution of the rv changes across both facet and x-axis variable."}

sim_varf_dist1 = function(nx, nfacet){
  rep(dist_normal(seq(mean,mean*nfacet, by = mean), sd), each = nx)}

p1 <- sim_panel(nx = 2, nfacet = 3, ntimes = 500) %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

simulated_data <- sim_panel(nx = 2,
          nfacet = 3,
          ntimes = 500,
          sim_dist = rep(dist_normal(seq(5, 15, 5), 5), each = 2))

p2 <- simulated_data %>% ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

p3 <- simulated_data %>% ggplot(aes(x = as.factor(id_facet), y = sim_data)) + facet_wrap(~id_x) + geom_boxplot()

p4 <- sim_panel(nx = 2,
          nfacet = 3,
          ntimes = 500,
          sim_dist = dist_normal(seq(5,30, 5), 5)) %>% 
  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot()

ggpubr::ggarrange(p1, p2, p3, p4, nrow = 2, ncol = 2,
                  common.legend = TRUE,
                  labels = c("a", "b", "c", "d"))

```

<!-- in the same group would be important to bring out different patterns of the data. -->

## Methodology

### Characterising distribution

Each of the data subsets in the data structure have multiple observations and may vary widely across different subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain. The set of observations corresponding to each combination is assumed to be a sample from an unknown probability density function.
While the whole population of observations has certain characteristics, we can typically never measure all of them. Often shape, central tendency, and variability are the common characteristics used to describe the distribution. Another way to describe the probability distribution is through quantiles. (Define quantiles here) <!--There are two broad approaches to quantile estimation, viz, parameteric and non-parameteric. The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data.--> Sample quantiles could be thought to estimate the population quantiles. But there are a large number of different definitions used for sample quantiles. The median-unbiased estimator is recommended (Rob's paper) because of its desirable properties of a quantile estimator and can be defined independently of the underlying distribution.

### Distance between distributions

<!-- One of the most important class of divergence is the f-divergence and includes measures like Kullback-Leibler divergence, Hellinger distance etc. The continuous version of f -divergence is given by -->
<!-- $$D_f(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
<!-- $f : [0,\infty) \rightarrow R \cup \{\infty\}$ is a continuous convex function, and $f(1) = 0$.  -->

The most common divergence measure between distributions is the Kullback-Leibler (KL) divergence[@Kullback1951-jy] introduced by Solomon Kullback and Richard Leibler in 1951. The KL divergence, denoted $D(p(x), q(x))$ is a non-symmetric measure of the difference between two probability distributions $p(x)$ and $q(x)$ and is interpreted as the amount of information lost when $q(x)$ is used to approximate $p(x)$. Although the KL divergence measures the “distance” between two distributions, it is not a distance measure since it is not symmetric and does not satisfy the triangle inequality. The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and it always has a finite value. The square root of the Jensen-Shannon divergence is a metric, often referred to as Jensen-Shannon distance. Other common measures of distance are Hellinger distance, total variation distance and Fisher information metric.

In the context of this paper, the pairwise distances between the distributions of the measured variable are computed through Jensen-Shannon distance (JSD) which is based on Kullback-Leibler divergence and is defined by,

$$JSD(P||Q) = \frac{1}{2}D(P||M) + \frac{1}{2}D(Q||M)$$
where $M = \frac{P+Q}{2}$ and 
$D(P||Q) := \int^\infty_{-\infty} p(x)f(\frac{p(x)}{q(x)})$ is the KL divergence between distributions $p(x)$ and $q(x)$. Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.

<!-- The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where, -->
<!-- $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.   -->

<!-- We call this measure of variation as  Median Maximum Pairwise Distances (MMPD). -->

<!-- #### Distribution of Jensen-Shannon distances -->

<!-- Jensen-Shannon distances (JSD) are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$. -->
<!-- With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->

### Definitions

Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let the pairwise distances between pairs $(a_{j} b_{k}, a_{j'}, b_{k})$ be denoted as $d_{(jk, j'k')} = JSD(a_{j}b_{k}, a_{j'}b_{k'})$. Pairwise distances could be within-facets or between-facets. Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined. Pairwise distances are within-facets ($d_{w}$) when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)). Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.

```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets. The dotted lines represent the distances between different categories. Panel 1) and 2) show the between-facet distances. Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively. When categories are ordered, distances should only be considered for consecutive x-axis categories. Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}
knitr::include_graphics(here::here("paper/Figs/dist_explain.png"))
```

From Section \ref{sec:idea}, the idea is to put more weights on within-facet distances than between-facet distances. Hence, for a suitable tuning parameter $\lambda > 1$, the pairwise distances $d_{(jk, j'k')}$ are transformed based on the distance type as follows:

\begin{equation}
d*_{(j,k), (j'k')} =
\begin{cases}
\lambda d_{(jk), (j'k')},& \text{if } d = d_w\\
(1-\lambda)  d_{(jk), (j'k')},              & \text{if } d = d_b\\
\end{cases}
\end{equation}

The maximum weighted pairwise distances are defined as:
$$WPD = max_{j, j', k, k'}(d*_{(jk), (j'k')})
\forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$.

<!-- _j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$. Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively. Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$. There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$). Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis. -->

 <!-- old -->
<!-- Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$. Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively. Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$. There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$). Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis.  -->

<!-- The algorithm employed for computing the distance measure is summarized as follows: -->

<!-- 1. Fix harmony pair $(C_i, C_j)$. -->

<!-- 2. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$. -->

<!-- 3. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories. -->

<!-- 4. Identify maximum within the $m$ computed distances. -->

<!-- <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.  --> -->

<!-- 5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$. -->

<!-- 6. Compute the distance measure MMPD_raw = median $(M_1, M_2, \dots, M_K)$. -->

<!-- \begin{algorithm}[!thb] -->
<!-- 	\caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}. -->
<!-- 	\begin{algorithmic}[1] -->
<!-- 		\Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}. -->
<!-- 		\For{$k=1:K$, $j=1:J$} -->
<!-- 		\State Find distances between pairs of all possible combinations of categories $(a_jb_k,a_j'b_k')$ by computing JSD between quantiles of the measured variable $q(v)$ across these combinations. -->
<!-- 		\State $d \gets JSD(\tilde{q(v)_{a_{j}b_{k}}},\tilde{q(v)_{a_{j}'b_{k}'}})$ -->
<!-- 		\If {$b_k = b_k'$} -->
<!-- 		\State $d* \gets \lambda d$  \Comment{upweight within-facet distances} -->
<!-- 		\Else -->
<!-- 		\State $d* \gets 1/\lambda d $ \Comment{downweight across-facet distances} -->
<!-- 		\EndFor -->
<!-- 		\State Set the raw distance measure as $max(d*)$ where max is taken over all $j, j', k, k'$. -->

<!-- 		\EndProcedure -->
<!-- 	\end{algorithmic} -->
<!-- \end{algorithm} -->

<!-- A stronger measure "max" is chosen for aggregating x-axis categories compared to "median" for aggregating facet categories as the measure is intended to put more importance in pointing towards distributional differences between x-axis categories than for facet categories. -->

@@

### Normal Distribution

```{r normal-diff, fig.cap="Ridge plots of raw wpd is shown for N(0,1) and N(5,1) distribution. The densities change across different facet and x levels but look same for the two distributions, which implies wpd value is unaffected by the change is mean value of the normal distribution"}
knitr::include_graphics(here::here("simulations/null_design/figs/normal_ridge_nxbynfacet_wpd_N01.png"))
```

### Gamma Distribution

```{r gamma-diff, fig.cap="Ridge plots of raw wpd is shown for G(0.5,1) and G(2,1) distribution. The densities change across different facet and x levels and also for the two distributions, which implies wpd value is affected by the change in location value when distributions are not normal."}
knitr::include_graphics(here::here("simulations/null_design/figs/gamma_ridge_nxbynfacet.png"))
```

<!-- ### Choice of alternate designs -->

<!-- A normal distribution is taken for this purpose. We would calibrate the values of $WPD$ for different designs illustrated in Section \ref{sec:idea}. -->

<!-- $$\mu_{j.} = \mu + j\omega$$ -->
<!-- or, $$\mu_{.k} = \mu + k\omega$$ where,    -->
<!-- $j$ and $k$ are the index of the x and facet category respectively  -->
<!-- $\mu$ is the mean of the first category and $\omega$ is the increment from the starting value. -->

<!-- mean = 0 -->
<!-- range_mean = c(0, 1, 3,  5, 7, 15, 20) -->
<!-- sd = 1 -->
<!-- range_sd = c(1, 3,  5, 7, 15, 20) -->
<!-- range_w =  c(1, 2, 3, 4, 5) -->
<!-- range_lambda = seq(0.1, 0.9, 0.05) -->

## Choice of tuning parameters and effect of different increments

Objective: Choice of optimal tuning parameter $lambda$ with the help of two alternate designs (vary_x and vary_f)  for different nx and nfacets. The lambda for which the two designs intersect have to be chosen to put appropriate weights for within-facet and between-facet distances.

Assumptions: The initial combination has a $N(\mu,\sigma)$ distribution and then every combination increases as follows:
- $$\mu_{j.} = \mu + j\omega$$ (for design vary_x)
- $$\mu_{.k} = \mu + k\omega$$ (for design vary_f)
Value of the parameters chosen for this simulation are $\mu = 0$, $\sigma = 1$ and $\omega = {1, 2, \dots, 10}$

Fixing omega, where the design interacts

```{r tuning-one-omega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```

How value of lambda changes with increasing omega

```{r tuning-parameter-intersection}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation?

```{r lambda-omega-mean}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_mean_3_2.png"))
```

How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean?

```{r lambda-omega-sd}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_sd_3_2.png"))
```

## Alternate designs

```{r lamba-omega-alternate-design}
knitr::include_graphics((here::here("simulations/result_report/wpd_alterate_designs.png")))
```

 Distribution across facet and x categories
<!-- ### Need for normalisation -->

From Figure \ref{fig:dist-across-facets} and \ref{fig:dist-across-x} shows the MMPD distribution is different for different levels of facets and x-axis levels. With increasing number of facets, the location of the distribution shift rightwards and with increasing x-axis levels, the scale of the distribution reduces.

Median and maximum distances are affected by the number of categories considered and hence the distance measure $MMPD_{raw}$, which is a combination of median and maximum would also be influenced by the number of levels. It would have higher values if $C_i$ or $C_j$ has higher levels. We would ideally want a higher value of the measure if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. In Figures \ref{fig:dist-across-facets} and \ref{fig:dist-across-x}, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which $MMPD_{raw}$ has been constructed $100$ times to compute the distribution. <!--More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.-->

Therefore, in order to compare $MMPD_{raw}$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure.

```{r hist-nx-nfacet, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/hist_nxbyfacet.png"))
```

```{r dist-across-facets, fig.cap = "Distribution of $MMPD_{raw}$ for different facet levels are shown. The location of the distribution shifts rightwards implying that as the number of facets increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained.", eval = FALSE}
here::here()
knitr::include_graphics(here::here("simulations/result_report","nx_by_facet_raw.png"))
```

<!-- # ```{r dist-across-facets, fig.cap = "Distribution of $MMPD_{raw}$ for different facet levels are shown. The location of the distribution shifts rightwards implying that as the number of facets increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained. "} -->
<!-- # mmpd_raw = read_rds("../hakear-drake/data/data_all23457914n.rds")  -->
<!-- # mmpd_xmore = read_rds("../hakear-drake/data/data_all213045x.rds")  -->
<!-- # mmpd_fmore = read_rds("../hakear-drake/data/data_all_213045f_allx.rds") -->
<!-- #  -->
<!-- # data_all = bind_rows(mmpd_raw, mmpd_xmore) -->
<!-- #  -->
<!-- # data_long = data_all %>% pivot_longer(c(1:2), names_to = "category", values_to = "levels") -->
<!-- #  -->
<!-- # g1 <- data_all %>%  -->
<!-- #   filter(!(nx %in% c(3, 4))) %>%  -->
<!-- #   ggplot() + -->
<!-- #   ggridges::geom_density_ridges(aes(x = raw_mmpd,  -->
<!-- #                                     y = as.factor(nfacet))) + -->
<!-- #   facet_wrap(~nx, nrow = 2) -->
<!-- # g1 -->
<!-- # ``` -->

```{r dist-across-x, fig.cap = "Distribution of $MMPD_{raw}$ for different x levels are shown. The scale of the distribution becomes less and less implying that as the number of x levels increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained as the spread of the distribution decreases. ", eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/","nfacet_by_nx_raw.png"))
```

>>>>

---
title: "vary_distribution"
author: "Sayani Gupta"
date: "18/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

### Standard normal distribution

```{r normal, fig.cap="Density plots of raw wpd is shown for N(0,1) distribution. The densities change across different facet and x levels, which implies wpd value changes with increasing and facet levels"}

knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/nxbyfacet_density_wpd_N01.png"))
```

### Comparing normal distributions with different means

```{r normal-diff, fig.cap="Ridge plots of raw wpd is shown for N(0,1) and N(5,1) distribution. The densities change across different facet and x levels but look same for the two distributions, which implies wpd value is unaffected by the change in mean value of the normal distribution"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_normal.png"))
```

### Gamma Distribution with different locations

```{r gamma-diff, fig.cap="Ridge plots of raw wpd is shown for G(0.5,1) and G(2,1) distribution. The densities change across different facet and x levels and also for the two distributions, which implies wpd value is affected by the change in location value when distributions are not normal."}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/extra/gamma_ridge_nxbynfacet.png"))
```

### Gamma Distribution with different locations after quantile transformation

```{r gamma-diff-trans, fig.cap="Ridge plots of raw wpd is shown for G(0.5,1) and G(2,1) distribution after quantile transformation looks similar and hence is unaffected by change in location"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_gamma.png"))
```
